The function--sheaf dictionary for smooth commutative group schemes over finite fields

### 2014-06-25 dictionary.tex
Fixed section on finite quasicharacter sheaves!
Also played with notation for rigid, finite, flabby, quasicharacter
sheaves. I’m not done with Aut(L) yet.

%%%%%%%%%%%% MISCELLANEOUS OPERATORS %%%%%%%%%%%%
\newcommand{\sheafHom}{{\mathscr{H}\hskip-4pt{\it o}\hskip-2pt{\it m}}}
\newcommand{\abs}[1]{{\vert #1 \vert}}
\newcommand{\ceq}{{\, :=\, }}
\newcommand{\tq}{{\ \vert\ }}
\newcommand{\iso}{{\ \cong\ }}
%% Limits
\newcommand{\invlim}[1]{\lim\limits_{\overleftarrow{#1}}}
\newcommand{\dirlim}[1]{\lim\limits_{\overrightarrow{#1}}}
\newcommand{\limit}[1]{\mathop{\textsc{lim}}\limits_{#1}}
\newcommand{\colimit}[1]{\mathop{\textsc{colim}}\limits_{#1}}
%% Fonts for quasicharacter sheaves
\newcommand{\qcs}[1]{{\mathcal{#1}}}
\newcommand{\gqcs}[1]{{\mathcal{\bar #1}}}
\newcommand{\dualgcs}[1]{\gqcs{#1}^\dagger}
\newcommand{\dualcs}[1]{\qcs{#1}^\dagger}
%% Categories
\newcommand{\lQC}{{\mathcal{Q\hskip-0.8pt C}}}
\newcommand{\lQCiso}[1]{\lQC(#1)_{/\text{iso}}}
\newcommand{\QC}{{\lQC_{\text{rig}}}}
\newcommand{\QCb}{{\lQC_0}}
\newcommand{\QCf}{{\lQC_f}}
\newcommand{\QCiso}[1]{\QC(#1)_{/\text{iso}}}
\newcommand{\QCbiso}[1]{\QCb(#1)_{/\text{iso}}}
\newcommand{\QCfiso}[1]{\QCf(#1)_{/\text{iso}}}
\newcommand{\trFrob}[1]{t_{#1}}
\newcommand{\TrFrob}[1]{T_{#1}}
%% amsmath replacements for \atop
\newcommand{\latop}[2]{\genfrac{}{}{0pt}{0}{#1}{#2}}
\newcommand{\partop}[2]{\genfrac{(}{)}{0pt}{0}{#1}{#2}}
%% Labeled items
\makeatletter
\newcommand{\labitem}[2]{
\def\@itemlabel{\textbf{#1}}
\item
\def\@currentlabel{#1}\label{#2}}
\makeatother
%% Shorthand for bars
\renewcommand{\bf}{\bar{f}}
\newcommand{\bg}{\bar{g}}
\newcommand{\bm}{\bar{m}}
\newcommand{\bG}{\bar{G}}
\newcommand{\bH}{\bar{H}}
%% Spacing control
\newcommand{\tight}[3]{\hspace{-#1pt}{#2}\hspace{-#3pt}}
\newcommand{\GxG}{\text{$G \tight{1}{\times}{1} G$}}
\newcommand{\bGxG}{\text{$\bar{G} \tight{1}{\times}{1} \bar{G}$}}
\newcommand{\bfxf}{\text{$\bar{f} \tight{1}{\times}{1} \bar{f}$}}
\newcommand{\GxxG}{\text{$G \tight{1}{\times}{1} G$}}
\newcommand{\LxL}{\text{$\gqcs{L} \tight{0}{\boxtimes}{0} \gqcs{L}$}}
\newcommand{\ExE}{\text{$\qcs{E}\tight{0}{\boxtimes}{0}\qcs{E}$}}
\newcommand{\bExE}{\text{$\gqcs{E}\tight{0}{\boxtimes}{0}\gqcs{E}$}}
\newcommand{\AxA}{\text{$A \tight{1}{\times}{1} A$}}
\newcommand{\BxB}{\text{$B \tight{1}{\times}{1} B$}}
\newcommand{\GzxGz}{\text{$G^0 \tight{1}{\times}{1} G^0$}}
%% Margin notes
\newcommand\Clifton[1]{\marginpar{\smaller\smaller CC: #1}}
\newcommand\David[1]{\marginpar{\smaller\smaller DR: #1}}
%% Hyphenation override
\hyphenation{quasi-character}

@@

Now let $G$ be any smooth commutative group scheme over $\Fq$.
Two new features must be addressed in order to geometrize characters of $G(\Fq)$.
First, we must pass from $\ell$-adic local systems on $G$ to Weil local systems on $G$
and, second, we must keep track of the isomorphisms \eqref{introbox}.
Both issues are addressed in the main definition of this paper:
A \emph{rigid quasicharacter sheaf} on $G$ is a triple $\qcs{L}\ceq
(\gqcs{L},\mu,\phi)$ where $\gqcs{L}$ is an $\ell$-adic local system on $\bG$ of rank~one and $\mu: m^*
\gqcs{L} \to \LxL$ and $\phi : \Frob{G}^*\gqcs{L} \to \gqcs{L}$ are isomorphisms satisfying natural compatibility conditions; here $\Frob{G}$ is the Frobenius automorphism of ${\bar G}$.
A \emph{quasicharacter sheaf} on $G$ is a Weil sheaf $(\gqcs{L},\phi)$ on $G$
for which $(\gqcs{L},\mu,\phi)$ is a rigid quasicharacter sheaf on $G$, for some $\mu$.
The difference between these two notions is detected by $\Hh^2(\pi_0({\bar G}),\EEx)$, where $\pi_0({\bar G})$ is the component group for ${\bar G}$.
%namely, when $\Hh^2(\pi_0({\bar G}),\EEx)$ is non-trivial, if $(\gqcs{L},\phi)$ is a quasicharacter sheaf on $G$ then there are many $\mu$ for which $(\gqcs{L},\mu,\phi)$ is a rigid quasicharacter sheaf on $G$;
%
 In the main result of this paper, Theorem~\ref{thm:snake},
we form a rigid monoidal category $\lQC(G)$ of quasicharacter sheaves on $G$
and show that the Trace of Frobenius provides a functorial group isomorphism
\begin{equation}
\lQCiso{G} \to \Hom(G(\Fq), \EEx).
\end{equation}

@@

We close the introduction with a brief description of the structure of this paper.
After setting some basic notation in Section~\ref{sec:notation},
in Section~\ref{sec:category} we define quasicharacter sheaves and give an interpretation of
them as Weil sheaves equipped with an isomorphism of the form \eqref{introbox}.
Section~\ref{sec:Frob} introduces Trace of Frobenius,
one direction of the function--sheaf dictionary.
We define pullbacks and products of quasicharacter sheaves in Section~\ref{sec:pullback}, and use
them to prove a functoriality result on Trace of Frobenius.
In Section~\ref{sec:basechange} we study how quasicharacter sheaves behave under base change
and Weil restriction.
Sections~\ref{sec:bounded} and \ref{sec:finite} describe bounded quasicharacter sheaves and finite quasicharacter sheaves, respectively,
forming two important subcategories of $\QC(G)$.
%These subcategories play a key role in the proof of Proposition~\ref{prop:restriction}.
In Section \ref{sec:connected} we show that all quasicharacter sheaves on $G$ are finite quasicharacter sheaves when $G$ is connected, and that Trace of Frobenius is an isomorphism in this case.
We shift to \'etale group schemes $G$ in Section~\ref{sec:etale} and find a cohomological interpretation
of quasicharacter sheaves in this case.
In Section~\ref{sec:restriction} we show that for non-connected $G$, restriction to the identity component
induces an essentially surjective functor of quasicharacter sheaves.
Section~\ref{sec:snake} presents the main theorem of this paper: Trace of Frobenius
is a functorial isomorphism from isomorphism classes of quasicharacter sheaves on $G$
to characters of $G(\Fq)$.

@@

Throughout this paper, $G$ is a smooth, commutative group scheme
over a finite field $\Fq$, and $m : \GxG\to G$ is its multiplication morphism.
%For the sake of brevity, we define a \emph{$\QC$-scheme} to be such a $G$ with the additional property that the geometric component group of $G$ is finitely generated.  This last condition is only needed in the proof of Lemma \ref{lemma:section}.  A morphism of $\QC$-schemes is a morphism of group schemes over $\Fq$.

@@

Morphisms in the category of quasicharacter sheaves on $G$, denoted by $\QC(G)$, are defined in the natural way:
\begin{enumerate}
\labitem{(QC.4)}{QC.4} if $\qcs{L} = (\gqcs{L},\mu,\phi)$ and
  $\qcs{L'} = (\gqcs{L'},\mu',\phi')$ are quasicharacter sheaves on $G$ then
  a morphism $\qcs{L} \to \qcs{L}'$ is a map $\alpha : \gqcs{L} \to \gqcs{L'}$
  of constructible $\ell$-adic sheaves on $\bG$ such that the following diagrams both commute.
  \[
  \begin{tikzcd}[column sep=40]
  \Frob{G}^* \gqcs{L} \rar{\Frob{G}^* \alpha} \arrow[swap]{d}{\phi} & \Frob{G}^* \gqcs{L'} \dar{\phi'}
  & & \arrow[swap]{d}{\mu} m^* \gqcs{L} \rar{m^* \alpha} & m^* \gqcs{L'} \dar{\mu'} \\
  \gqcs{L} \rar{\alpha} & \gqcs{L'}
  & {} & \LxL \rar{\tight{1}{\alpha\boxtimes \alpha}{1}} & \gqcs{L'} \tight{0}{\boxtimes}{0} \gqcs{L'}
  \end{tikzcd}
  \]
\end{enumerate}

The rule $(\gqcs{L},\mu,\phi) \mapsto (\gqcs{L},\phi)$ defines a forgetful functor from quasicharacter
sheaves on $G$ to ($\ell$-adic) Weil sheaves on $G$ \cite{deligne:80a}*{Def.~1.1.10 (i)}.

A Weil sheaf $(\gqcs{L},\phi)$ on $G$
may be interpreted as a constructible $\ell$-adic sheaf $\gqcs{L}$ on $\bG$ together with
an action of the Weil group $\Weil{}$ on $\gqcs{L}$ compatible with the action of
$\Gal(\bFq/\Fq)$ on $\bG$. We now briefly review this point of view for later use in
Section~\ref{sec:basechange}; see also \cite{deligne:80a}*{Def.~1.1.10}.  If $w \ceq \Frob{}^n$ we write $w_G$ for $\Frob{G}^n$.
For each such $w$ we define an isomorphism $\varphi(w) : w_G^* \gqcs{L}\to \gqcs{L}$ by
$
\varphi(w) \ceq  \phi \circ \Frob{G}^*(\phi) \circ \cdots \circ (\Frob{G}^{n-1})^*(\phi)
$;
these isomorphisms satisfy $\varphi(uv) = \varphi(v) \circ v_G^* \varphi(u)$ for $u,v\in \Weil{}$, and
we have $\phi = \varphi(\Frob{})$.
If we further define $\varphi_1(w) \ceq (w_G)_*(\varphi(w)^{-1})$ then
$\varphi_1(w) : (w_G)_* \gqcs{L}\to \gqcs{L}$ is an isomorphism and
$\varphi_1(uv) = \varphi_1(u) \circ (u_G)_* \varphi_1(v)$ for $u,v\in \Weil{}$.
Thus the pair $(\gqcs{L},\varphi_1)$ almost satisfies the criteria for an action of $\Weil{}$
on $\gqcs{L}$ compatible with the action of $\Gal(\bFq/\Fq)$ in the sense of
\cite{deligne:SGA4.5}*{Expos\'e XIII, 1.1}, failing only because $\Weil{}$ is not profinite.
As we will see in Corollary~\ref{cor:bounded-and-finite}, our use of $\Weil{}$ rather than $\Gal(\bFq/\Fq)$
allows quasicharacter sheaves to correspond to unbounded characters.

The category $\QC(G)$ of quasicharacter sheaves on $G$ is a rigid monoidal category
\cite{etingof:09a}*{\S1.10} under the tensor product
$\qcs{L} \otimes \qcs{L'} \ceq (\gqcs{L}\otimes\gqcs{L'}, \mu\otimes\mu', \phi\otimes \phi')$,
with duals given by applying the sheaf hom functor
$\sheafHom(\ - \ ,\EE)$.
The category of quasicharacter sheaves on $G$ is not abelian, so $\QC(G)$
is not a tensor category in the sense of \cite{deligne:02a}*{0.1}.
This rigid monoidal category structure for $\QC(G)$ gives the set $\QCiso{G}$
of isomorphism classes in $\QC(G)$ the structure of a group.

@@

We now begin developing the tools to relate $\QCiso{G}$ to $G(\Fq)^*$.
%
Let $(\gqcs{L},\phi)$ be a Weil sheaf on $G$. Every $g\in G(\Fq)$
determines a point $\bg$ on $\bG$ fixed by $\Frob{G}$ and
therefore an automorphism $\phi_{\bg}$ of $\gqcs{L}_{\bg}$. Set
\begin{equation}\label{trWeil}
%\begin{equation}\label{trWeil}
\trFrob{(\gqcs{L},\phi)}(g) \ceq \trace(\phi_{\bg}, \gqcs{L}_{\bg})
%\end{equation}
\end{equation}
and note that if $(\gqcs{L},\phi) \iso (\gqcs{L'},\phi')$ then $\trFrob{(\gqcs{L},\phi)} = \trFrob{(\gqcs{L'},\phi')}$.
 If $\qcs{L} = (\gqcs{L},\mu,\phi)$ is a quasicharacter sheaf
then we will abbreviate $\trFrob{(\gqcs{L},\phi)}$ to $\trFrob{\qcs{L}}$.
%\begin{definition}\label{def:trFrob}
We will refer to the function
\begin{align*}
\TrFrob{G}: \QCiso{G} &\to G(\Fq)^* \\
\qcs{L} &\mapsto t_{\qcs{L}}: g\mapsto \trace(\phi_{\bg}, \gqcs{L}_{\bg})
\end{align*}
as the \emph{Trace of Frobenius} for $G$.
%\todo{Does this really warrant a declared definition?
%Instead, we can refer to this very short section, when needed.}
%\end{definition}

The isomorphism $\mu : \bm^* \gqcs{L} \to \LxL$ together with
\ref{QC.3} guarantee
that the function $\trFrob{\qcs{L}} : G(\Fq)\to \EEx$ is a group homomorphism.  If we write
$\trFrob{\qcs{L}} \cdot \trFrob{\qcs{L'}}$ for pointwise multiplication, then
 $\trFrob{\qcs{L}\otimes \qcs{L'}} = \trFrob{\qcs{L}} \cdot \trFrob{\qcs{L'}}$.
 Therefore $\TrFrob{G}$ is a homomorphism, which is functorial in the following sense:

\begin{proposition}\label{prop:functorialG}
Trace of Frobenius $\TrFrob{G} : \QCiso{G} \to G(\Fq)^*$ is a natural transformation
between the two contravariant additive functors
\begin{align*}
F_1 : G &\mapsto \QCiso{G} \\
F_2 : G &\mapsto G(\Fq)^*
\end{align*}
from the category of smooth commutative group schemes over $\Fq$ to the category of abelian groups.
\end{proposition}

@@

\begin{proposition}\label{prop:pullback}
  If $f : H\to G$ is a morphism of smooth commutative group schemes over $\Fq$ then
  \begin{align*}
  f^* : \QC(G) &\to \QC(H) \\
  (\gqcs{L},\mu,\phi) &\mapsto (\bf^*\gqcs{L},(\bfxf)^*\mu,\bf^*F)
  \end{align*}
  defines a monoidal functor dual to $f \colon H(\Fq) \to G(\Fq)$:
  \[
  \begin{tikzcd}[row sep=20, column sep=30]
   \QCiso{G} \rar{f^*} \arrow[swap]{d}{\TrFrob{G}} & \QCiso{H} \dar{\TrFrob{H}} \\
   G(\Fq)^* \rar & H(\Fq)^*
  \end{tikzcd}
  \]
  is a commutative diagram of groups.  Moreover, $(f\circ g)^* = g^* \circ f^*$.
\end{proposition}

@@

We end this section with a simple result on products of quasicharacter sheaves,
which is needed to show that the isomorphism above is functorial.

>>>>

%%%%%%%%%%%% MISCELLANEOUS OPERATORS %%%%%%%%%%%%
\newcommand{\sheafHom}{{\mathscr{H}\hskip-4pt{\it o}\hskip-2pt{\it m}}}
\newcommand{\abs}[1]{{\vert #1 \vert}}
\newcommand{\ceq}{{\, :=\, }}
\newcommand{\tq}{{\ \vert\ }}
\newcommand{\iso}{{\ \cong\ }}
%% Limits
\newcommand{\invlim}[1]{\lim\limits_{\overleftarrow{#1}}}
\newcommand{\dirlim}[1]{\lim\limits_{\overrightarrow{#1}}}
\newcommand{\limit}[1]{\mathop{\textsc{lim}}\limits_{#1}}
\newcommand{\colimit}[1]{\mathop{\textsc{colim}}\limits_{#1}}
%% Fonts for quasicharacter sheaves
\newcommand{\qcs}[1]{{\mathcal{#1}}}
\newcommand{\gqcs}[1]{{\mathcal{\bar #1}}}
\newcommand{\dualgcs}[1]{\gqcs{#1}^\dagger}
\newcommand{\dualcs}[1]{\qcs{#1}^\dagger}
%% Categories
\newcommand{\QC}{{\mathcal{Q\hskip-0.8pt C}}}
\newcommand{\lQC}{{\QC_\text{flb}}}
\newcommand{\lQCiso}[1]{\lQC(#1)_{/\text{iso}}}
\newcommand{\rQC}{{\QC_{\text{rig}}}}
\newcommand{\rbQC}{{\QC_0}}
\newcommand{\rfQC}{{\QC_\text{fin}}}
\newcommand{\rQCiso}[1]{\rQC(#1)_{/\text{iso}}}
\newcommand{\rbQCiso}[1]{\rbQC(#1)_{/\text{iso}}}
\newcommand{\rfQCiso}[1]{\rfQC(#1)_{/\text{iso}}}
\newcommand{\trFrob}[1]{t_{#1}}
\newcommand{\TrFrob}[1]{T_{#1}}
%% amsmath replacements for \atop
\newcommand{\latop}[2]{\genfrac{}{}{0pt}{0}{#1}{#2}}
\newcommand{\partop}[2]{\genfrac{(}{)}{0pt}{0}{#1}{#2}}
%% Labeled items
\makeatletter
\newcommand{\labitem}[2]{
\def\@itemlabel{\textbf{#1}}
\item
\def\@currentlabel{#1}\label{#2}}
\makeatother
%% Shorthand for bars
\renewcommand{\bf}{\bar{f}}
\newcommand{\bg}{\bar{g}}
\newcommand{\bm}{\bar{m}}
\newcommand{\bG}{\bar{G}}
\newcommand{\bH}{\bar{H}}
%% Spacing control
\newcommand{\tight}[3]{\hspace{-#1pt}{#2}\hspace{-#3pt}}
\newcommand{\GxG}{\text{$G \tight{1}{\times}{1} G$}}
\newcommand{\bGxG}{\text{$\bar{G} \tight{1}{\times}{1} \bar{G}$}}
\newcommand{\bfxf}{\text{$\bar{f} \tight{1}{\times}{1} \bar{f}$}}
\newcommand{\GxxG}{\text{$G \tight{1}{\times}{1} G$}}
\newcommand{\LxL}{\text{$\gqcs{L} \tight{0}{\boxtimes}{0} \gqcs{L}$}}
\newcommand{\ExE}{\text{$\qcs{E}\tight{0}{\boxtimes}{0}\qcs{E}$}}
\newcommand{\bExE}{\text{$\gqcs{E}\tight{0}{\boxtimes}{0}\gqcs{E}$}}
\newcommand{\AxA}{\text{$A \tight{1}{\times}{1} A$}}
\newcommand{\BxB}{\text{$B \tight{1}{\times}{1} B$}}
\newcommand{\GzxGz}{\text{$G^0 \tight{1}{\times}{1} G^0$}}
%% Margin notes
\newcommand\Clifton[1]{\marginpar{\smaller\smaller CC: #1}}
\newcommand\David[1]{\marginpar{\smaller\smaller DR: #1}}
%% Hyphenation override
\hyphenation{quasi-character}

@@

Now let $G$ be any smooth commutative group scheme over $\Fq$.
Two new features must be addressed in order to geometrize characters of $G(\Fq)$.
First, we must pass from $\ell$-adic local systems on $G$ to Weil local systems on $G$
and, second, we must keep track of the isomorphisms \eqref{introbox}.
Both issues are addressed in the main definition of this paper:
A \emph{rigid quasicharacter sheaf} on $G$ is a triple $\qcs{L}\ceq
(\gqcs{L},\mu,\phi)$ where $\gqcs{L}$ is an $\ell$-adic local system on $\bG$ of rank~one and $\mu: m^*
\gqcs{L} \to \LxL$ and $\phi : \Frob{G}^*\gqcs{L} \to \gqcs{L}$ are isomorphisms satisfying natural compatibility conditions; here $\Frob{G}$ is the Frobenius automorphism of ${\bar G}$.
A \emph{flabby quasicharacter sheaf} on $G$ is a Weil sheaf $(\gqcs{L},\phi)$ on $G$
for which $(\gqcs{L},\mu,\phi)$ is a rigid quasicharacter sheaf on $G$, for some $\mu$.
The difference between these two notions is detected by $\Hh^2(\pi_0({\bar G}),\EEx)$, where $\pi_0({\bar G})$ is the component group for ${\bar G}$.
%namely, when $\Hh^2(\pi_0({\bar G}),\EEx)$ is non-trivial, if $(\gqcs{L},\phi)$ is a quasicharacter sheaf on $G$ then there are many $\mu$ for which $(\gqcs{L},\mu,\phi)$ is a rigid quasicharacter sheaf on $G$;
%
 In the main result of this paper, Theorem~\ref{thm:snake},
we form a rigid monoidal category $\lQC(G)$ of flabby quasicharacter sheaves on $G$
and show that the Trace of Frobenius provides a functorial group isomorphism
\begin{equation}
\lQCiso{G} \to \Hom(G(\Fq), \EEx).
\end{equation}

@@

We close the introduction with a brief description of the structure of this paper.
After setting some basic notation in Section~\ref{sec:notation},
in Section~\ref{sec:category} we define rigid quasicharacter sheaves and give an interpretation of
them as Weil sheaves equipped with an isomorphism of the form \eqref{introbox}.
Section~\ref{sec:Frob} introduces Trace of Frobenius,
one direction of the function--sheaf dictionary.
We define pullbacks and products of rigid quasicharacter sheaves in Section~\ref{sec:pullback}, and use
them to prove a functoriality result on Trace of Frobenius.
In Section~\ref{sec:basechange} we study how rigid quasicharacter sheaves behave under base change
and Weil restriction.
Sections~\ref{sec:bounded} and \ref{sec:finite} describe bounded quasicharacter sheaves and finite quasicharacter sheaves, respectively,
forming two important subcategories of $\rQC(G)$.
%These subcategories play a key role in the proof of Proposition~\ref{prop:restriction}.
In Section \ref{sec:connected} we show that all rigid quasicharacter sheaves on $G$ are finite quasicharacter sheaves when $G$ is connected, and that Trace of Frobenius is an isomorphism in this case.
We shift to \'etale group schemes $G$ in Section~\ref{sec:etale} and find a cohomological interpretation
of rigid quasicharacter sheaves in this case.
In Section~\ref{sec:restriction} we show that for non-connected $G$, restriction to the identity component
induces an essentially surjective functor of rigid quasicharacter sheaves.
Section~\ref{sec:snake} presents the main theorem of this paper: Trace of Frobenius
is a functorial isomorphism from isomorphism classes of flabby quasicharacter sheaves on $G$
to characters of $G(\Fq)$.

Comments about Joseph Bernstein's comments?

@@

Throughout this paper, $G$ is a smooth, commutative group scheme
over a finite field $\Fq$, and $m : \GxG\to G$ is its multiplication morphism.
%For the sake of brevity, we define a \emph{$\rQC$-scheme} to be such a $G$ with the additional property that the geometric component group of $G$ is finitely generated.  This last condition is only needed in the proof of Lemma \ref{lemma:section}.  A morphism of $\rQC$-schemes is a morphism of group schemes over $\Fq$.

@@

Morphisms in the category of rigid quasicharacter sheaves on $G$, denoted by $\rQC(G)$, are defined in the natural way:
\begin{enumerate}
\labitem{(QC.4)}{QC.4} if $\qcs{L} = (\gqcs{L},\mu,\phi)$ and
  $\qcs{L'} = (\gqcs{L'},\mu',\phi')$ are rigid quasicharacter sheaves on $G$ then
  a morphism $\qcs{L} \to \qcs{L}'$ is a map $\alpha : \gqcs{L} \to \gqcs{L'}$
  of constructible $\ell$-adic sheaves on $\bG$ such that the following diagrams both commute.
  \[
  \begin{tikzcd}[column sep=40]
  \Frob{G}^* \gqcs{L} \rar{\Frob{G}^* \alpha} \arrow[swap]{d}{\phi} & \Frob{G}^* \gqcs{L'} \dar{\phi'}
  & & \arrow[swap]{d}{\mu} m^* \gqcs{L} \rar{m^* \alpha} & m^* \gqcs{L'} \dar{\mu'} \\
  \gqcs{L} \rar{\alpha} & \gqcs{L'}
  & {} & \LxL \rar{\tight{1}{\alpha\boxtimes \alpha}{1}} & \gqcs{L'} \tight{0}{\boxtimes}{0} \gqcs{L'}
  \end{tikzcd}
  \]
\end{enumerate}

The category $\rQC(G)$ of rigid quasicharacter sheaves on $G$ is a rigid monoidal category
\cite{etingof:09a}*{\S1.10} under the tensor product
$\qcs{L} \otimes \qcs{L'} \ceq (\gqcs{L}\otimes\gqcs{L'}, \mu\otimes\mu', \phi\otimes \phi')$,
with duals given by applying the sheaf hom functor
$\sheafHom(\ - \ ,\EE)$.
The category of rigid quasicharacter sheaves on $G$ is not abelian, so $\rQC(G)$
is not a tensor category in the sense of \cite{deligne:02a}*{0.1}.
This rigid monoidal category structure for $\rQC(G)$ gives the set $\rQCiso{G}$
of isomorphism classes in $\rQC(G)$ the structure of a group.

\subsection{Flabby quasicharacter sheaves}\label{sec:loose}

Observe that the rule $(\gqcs{L},\mu,\phi) \mapsto (\gqcs{L},\phi)$ defines a forgetful functor from rigid quasicharacter
sheaves on $G$ to ($\ell$-adic) Weil sheaves on $G$ \cite{deligne:80a}*{Def.~1.1.10 (i)}.
%

%A Weil sheaf $(\gqcs{L},\phi)$ on $G$ may be interpreted as a constructible $\ell$-adic sheaf $\gqcs{L}$ on $\bG$ together with an action of the Weil group $\Weil{}$ on $\gqcs{L}$ compatible with the action of $\Gal(\bFq/\Fq)$ on $\bG$. 
%We now briefly review this point of view for later use in Section~\ref{sec:basechange}; see also \cite{deligne:80a}*{Def.~1.1.10}.  If $w \ceq \Frob{}^n$ we write $w_G$ for $\Frob{G}^n$.
%For each such $w$ we define an isomorphism $\varphi(w) : w_G^* \gqcs{L}\to \gqcs{L}$ by $\varphi(w) \ceq  \phi \circ \Frob{G}^*(\phi) \circ \cdots \circ (\Frob{G}^{n-1})^*(\phi);$
%these isomorphisms satisfy $\varphi(uv) = \varphi(v) \circ v_G^* \varphi(u)$ for $u,v\in \Weil{}$, and we have $\phi = \varphi(\Frob{})$.
%If we further define $\varphi_1(w) \ceq (w_G)_*(\varphi(w)^{-1})$ then $\varphi_1(w) : (w_G)_* \gqcs{L}\to \gqcs{L}$ is an isomorphism and $\varphi_1(uv) = \varphi_1(u) \circ (u_G)_* \varphi_1(v)$ for $u,v\in \Weil{}$.
%Thus the pair $(\gqcs{L},\varphi_1)$ almost satisfies the criteria for an action of $\Weil{}$ on $\gqcs{L}$ compatible with the action of $\Gal(\bFq/\Fq)$ in the sense of\cite{deligne:SGA4.5}*{Expos\'e XIII, 1.1}, failing only because $\Weil{}$ is not profinite.
%As we will see in Corollary~\ref{cor:bounded-and-finite}, our use of $\Weil{}$ rather than $\Gal(\bFq/\Fq)$ allows quasicharacter sheaves to correspond to unbounded characters.

\begin{definition}\label{def:lQC}
Let $G$ be a smooth commutative group scheme over $\Fq$.
A \emph{flabby quasicharacter sheaf} on $G$ is a Weil sheaf $(\gqcs{L},\phi)$ on $G$
for which $(\gqcs{L},\mu,\phi)$ is a rigid quasicharacter sheaf on $G$
for some $\mu: \bm^* \gqcs{L} \to \gqcs{L} \boxtimes \gqcs{L}$.
%The resulting objects lie in the essential image of the forgetful functor $(\gqcs{L},\mu,\phi) \mapsto (\gqcs{L},\phi)$ from $\rQC(G)$ to Weil sheaves on $G$.
We form the category of flabby quasicharacter sheaves by declaring that a morphism $(\gqcs{L},\phi) \to (\gqcs{L'},\phi')$ is a sheaf homomorphism $\alpha : \gqcs{L} \to \gqcs{L'}$ for which the two diagrams in Condition~\ref{QC.4} commute, for some choice of $\mu$, $\mu'$ such that $(\gqcs{L},\mu,\phi)$ and $(\gqcs{L'},\mu',\phi')$ are rigid quasicharacter sheaves on $G$.
The category of flabby quasicharacter sheaves on $G$ is denoted by $\lQC(G)$.
\end{definition}

Since objects in $\lQC(G)$ are Weil sheaves and since in the first diagram in Condition~\ref{QC.4} ensures that a morphism in $\lQC(G)$ is a morphism of Weil sheaves, it follows that $\lQC(G)$ is a subcategory of Weil sheaves on $G$.
%
Like $\rQC(G)$, the category $\lQC(G)$ of quasicharacter sheaves on $G$ is rigid monoidal. The functor $\rQC(G) \to \lQC(G)$ is monoidal and essentially surjective but neither full nor faithful, generally.
%

@@

We now introduce a tool which will help us understand isomorphism classes of objects in $\rQCiso{G}$ and $\lQCiso{G}$.

Let $(\gqcs{L},\phi)$ be a Weil sheaf on $G$. Every $g\in G(\Fq)$
determines a point $\bg$ on $\bG$ fixed by $\Frob{G}$. 
Together with the canonical isomorphism $(\Frob{G}^*\gqcs{L})_{\bg} \iso  \gqcs{L}_{\Frob{G}(\bg)}$, in this way the automorphism $\phi$ determines an automorphism $\phi_{\bg}$ of the $\EE$-vector space $\gqcs{L}_{\bg}$; let $\trace(\phi_{\bg};\gqcs{L}_{\bg})$ be the trace of $\phi_{\bg} \in \Aut_{\EE}(\gqcs{L}_{\bg})$ and let $\trFrob{(\gqcs{L},\phi)} : G(\Fq)\to \EE$ be the function defined by 
\begin{equation}\label{trWeil}
%\begin{equation}\label{trWeil}
\trFrob{(\gqcs{L},\phi)}(g) \ceq \trace(\phi_{\bg};\gqcs{L}_{\bg}),
%\end{equation}
\end{equation}
commonly called the {\em trace of Frobenius of $(\gqcs{L},\phi)$}.
Note that if $(\gqcs{L},\phi) \iso (\gqcs{L'},\phi')$ as Weil sheaves, 
then $\trFrob{(\gqcs{L},\phi)} = \trFrob{(\gqcs{L'},\phi')}$ as functions on $G(\Fq)$.

Now suppose $\qcs{L} = (\gqcs{L},\mu,\phi)$ is a rigid quasicharacter sheaf on $G$.
The isomorphism $\mu : \bm^* \gqcs{L} \to \LxL$ together with
\ref{QC.3} guarantee
that the function $\trFrob{(\qcs{L},\phi)} : G(\Fq)\to \EEx$ is a group homomorphism.  
Moreover, this group homomorphism is insensitive to $\phi$ in the sense 
that if $(\gqcs{L},\mu,\phi)$ and $(\gqcs{L},\mu,\phi')$ are rigid quasicharacter sheaves, then $\trFrob{(\qcs{L},\phi)} = \trFrob{(\qcs{L},\phi')}$. \todo{Sketch proof?}
This justifies abbreviating $\trFrob{(\gqcs{L},\phi)}$ to $\trFrob{\qcs{L}}$ in this case.
The diagram below defines two group homomorphisms, both denoted $\TrFrob{G}$ and both called the \emph{Trace of Frobenius} for $G$; we trust that context will remove any ambiguity regarding this notation.
\[
\begin{tikzcd}
\rQCiso{G} \arrow{rr} \arrow{dr}[swap]{\qcs{L} \mapsto t_{\qcs{L}}}{\TrFrob{G}} && \arrow{dl}[swap]{\TrFrob{G}}{(\gqcs{L},\phi) \mapsto t_{\gqcs{L},\phi}} \lQCiso{G} \\
& G(\Fq)^* & 
\end{tikzcd}
\]
As remarked above, the group homomorphism $\rQCiso{G} \to \lQCiso{G}$ is surjective;
we will see that $\rQCiso{G} \to \lQCiso{G}$ need not be injective.
However, in Theorem~\ref{thm:snake}, we will see that the group homomorphism $\rQCiso{G} \to G(\Fq)^*$ is surjective
while the main result of this paper, Theorem~\ref{thm:snake}, 
shows that $\lQCiso{G} \to G(\Fq)^*$ is an isomorphism.

\begin{proposition}\label{prop:functorialG}
Trace of Frobenius $\TrFrob{G} : \rQCiso{G} \to G(\Fq)^*$ is a natural transformation
between the two contravariant additive functors
\begin{align*}
F_1 : G &\mapsto \rQCiso{G} \\
F_2 : G &\mapsto G(\Fq)^*
\end{align*}
from the category of smooth commutative group schemes over $\Fq$ to the category of abelian groups.
Likewise, for $\TrFrob{G} : \lQCiso{G} \to G(\Fq)^*$, {\it mut.\hskip-1pt mut.}
\end{proposition}

@@

\begin{proposition}\label{prop:pullback}
  If $f : H\to G$ is a morphism of smooth commutative group schemes over $\Fq$ then
  \begin{align*}
  f^* : \rQC(G) &\to \rQC(H) \\
  (\gqcs{L},\mu,\phi) &\mapsto (\bf^*\gqcs{L},(\bfxf)^*\mu,\bf^*F)
  \end{align*}
  defines a monoidal functor dual to $f \colon H(\Fq) \to G(\Fq)$:
  \[
  \begin{tikzcd}[row sep=20, column sep=30]
   \rQCiso{G} \rar{f^*} \arrow[swap]{d}{\TrFrob{G}} & \rQCiso{H} \dar{\TrFrob{H}} \\
   G(\Fq)^* \rar & H(\Fq)^*
  \end{tikzcd}
  \]
  is a commutative diagram of groups.  Moreover, $(f\circ g)^* = g^* \circ f^*$.
\end{proposition}

@@

We end this section with a simple result on products of rigid quasicharacter sheaves,
which is needed to show that the isomorphism above is functorial.
